Reasoning about a Capability Machine with Local Capabilities\\ Provably Safe Stack and Return Pointer Management (without OS Support).

### 2017-02-09 csf-2017/paper.tex
lr pass

\begin{align*}
  &\memSat{\ms}{W}
    \text{ iff }
    \left\{\begin{aligned}
        &\exists P : \activeReg{W} \rightarrow \HeapSegments \ldotp \\
        &\quad (\hs = \biguplus_{r\in\activeReg{W}}P(r) \text{\footnotesize{ and }} \\
        &\quad \forall r \in \activeReg{W} \ldotp \exists H,s \ldotp\\
        &\qquad W(r) = (\_,s,\_,\_,H) \text{\footnotesize{ and }} \\
        &\qquad \npair[n]{P(r)} \in H(s)(\xi^{-1}(W)))\\
      \end{aligned}\right.
\end{align*}

A memory satisfies a world, written $\memSat{\ms}{W}$, 
if it can be partitioned into disjoint parts such that each part is
in the interpretation of an active (permanent or temporary) region.

@@

The \emph{expression relation}, $\stder$, defines what words are
capability safe when used for the program counter. A word is
capability safe as a program counter when it can be plugged into a
capability safe register file (i.e., a register file in $\stdrr$) and
paired with a memory satisfying the world to become a capability safe
configuration.

@@

The conditions $\entryCond{}$ and $\execCond{}$ are very similar and
differ only slightly. Executable capabilities can be updated to point
anywhere in memory, so they must be capability safe as a program
counter no matter where they point in memory. In contrast, enter
\lars{``capability safe as a program counter'' is a bit hard to
  understand - why not say that this is a capability one can jump to ?}
\lau{ When jumping to a capability, it becomes the program counter. A program counter is capability safe when it is the $\stder$ relation. When I say ``capability safe as a program counter'' I take it to mean being in the $\stder$-relation. }
capabilities are opaque and must be used from the address they point
to. They change permission when jumped to, so we require them to be
capability safe as a program counter when the permission is changed to
read/execute.
In both conditions, it must be safe to jump to the capability. This
means that the capability must be capability safe as a
program counter in certain future worlds. If the capability is global,
then it should be capability safe as a program counter in any private
future world. This means that it can only depend on permanent regions
because temporary regions may be revoked in a private future
world. On the other hand, local capabilities must only be capability
safe in public future worlds because they may depend on temporary
regions, so we cannot expect them to function correctly as program
counters if temporary regions are revoked.

@@

With the logical relation defined, we can now state a strong theorem
about the safety properties of the capability machine. Namely that any
capability that does not grant unreasonable authority is capability
safe as a program counter.
\begin{theorem}[Fundamental Theorem]
  \label{thm:ftlr}
  For all $n$, $\perm$, $\start$, $\addrend$, $\addr$, $\gl$, $W$, if one of the following holds:
  \begin{align*}
      & \bullet
        \begin{aligned}[t]
        &\perm = \exec \text{\footnotesize{ and }}\\
        &\npair{(\start,\addrend)} \in \readCond{}(\gl)(W)
      \end{aligned} \\
    & \bullet 
      \begin{aligned}[t]
        &\perm = \rwx \text{\footnotesize{ and }} \\
        &\npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{\footnotesize{ and }}\\
        &\npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\gl)(W)
      \end{aligned} \\
    & \bullet 
      \begin{aligned}[t]
        &\perm = \rwlx \text{\footnotesize{ and }}\\
        &\npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{\footnotesize{ and }}\\
        &\npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W),
      \end{aligned}
  \end{align*}
  then
  \[
    \npair{((\perm,\gl),\start,\addrend,\addr)} \in \stder(W)
  \]
\end{theorem}
The Fundamental Theorem allows us to prove capability safety for
untrusted programs. If we can show that the authority the capability
for the program grants is reasonable, then it is safe to
execute. Intuitively this is because the logical relation captures the
safe behaviour of the capability machine that do not allow you to
produce something capability \emph{unsafe} from something capability
safe. The Fundamental Theorem also serves as a sanity check of the
logical relation as it makes sure that a large class of capabilities
actually inhabits it.

@@

\bibliographystyle{IEEEtran}
\bibliography{IEEEabrv,references}
\end{document}
\grid

>>>>

\subsubsection{World satisfaction}
A memory satisfies a world, written $\memSat{\ms}{W}$, 
if it can be partitioned into disjoint parts such that each part is
in the interpretation of an active (permanent or temporary) region. 
\begin{align*}
  &\memSat{\ms}{W}
    \text{ iff }
    \left\{\begin{aligned}
        &\exists P : \activeReg{W} \rightarrow \HeapSegments \ldotp \\
        &\quad (\hs = \biguplus_{r\in\activeReg{W}}P(r) \text{\footnotesize{ and }} \\
        &\quad \forall r \in \activeReg{W} \ldotp \exists H,s \ldotp\\
        &\qquad W(r) = (\_,s,\_,\_,H) \text{\footnotesize{ and }} \\
        &\qquad \npair[n]{P(r)} \in H(s)(\xi^{-1}(W)))\\
      \end{aligned}\right.
\end{align*}

@@

The \emph{expression relation}, $\stder$, defines when a word is
capability safe as a program counter. That is the case when the word can be plugged into a
capability safe register file (i.e., a register file in $\stdrr$) and
paired with a memory satisfying the world to become a capability safe
configuration.

@@

The conditions $\entryCond{}$ and $\execCond{}$ are very similar and
differ only slightly. Executable capabilities can be updated to point
anywhere in memory, so they must be capability safe as a program
counter (in the $\stder$-relation) no matter where they point in memory. In contrast, enter
capabilities are opaque and must be used from the address they point
to. They change permission when jumped to, so we require them to be
capability safe as a program counter when the permission is changed to
read/execute.
In both cases, it must be safe to jump to the capability. This
means that the capability must be capability safe as a
program counter in a future world. If the capability is global,
then it should only depend on permanent regions and therefore
we require that it should be capability safe as a program counter in any \emph{private}
future world (recall that temporary regions may be revoked in a private future
world). On the other hand, a local capability should be capability
safe in any \emph{public} future world because it may depend on temporary
regions, so we cannot expect them to function correctly as program
counters if temporary regions are revoked.

@@

With the logical relation defined, we can now state a strong theorem
about the safety properties of the capability machine:
any capability that only grants reasonable authority is capability
safe as a program counter.
\begin{theorem}[Fundamental Theorem]
  \label{thm:ftlr}
  For all $n$, $\perm$, $\start$, $\addrend$, $\addr$, $\gl$, $W$, if one of the following holds:
  \begin{align*}
      & \bullet
        \begin{aligned}[t]
        &\perm = \exec \text{\footnotesize{ and }}\\
        &\npair{(\start,\addrend)} \in \readCond{}(\gl)(W)
      \end{aligned} \\
    & \bullet 
      \begin{aligned}[t]
        &\perm = \rwx \text{\footnotesize{ and }} \\
        &\npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{\footnotesize{ and }}\\
        &\npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\gl)(W)
      \end{aligned} \\
    & \bullet 
      \begin{aligned}[t]
        &\perm = \rwlx \text{\footnotesize{ and }}\\
        &\npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{\footnotesize{ and }}\\
        &\npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W),
      \end{aligned}
  \end{align*}
  then
  \[
    \npair{((\perm,\gl),\start,\addrend,\addr)} \in \stder(W)
  \]
\end{theorem}
The Fundamental Theorem allows us to prove capability safety for
untrusted programs: if the authority granted by
the capabilty for the untrustred program is reasonable, then it is safe to
execute. Intuitively, this is because our logical relation captures the
safe behaviours of the capability machine that do not allow you to
produce something \emph{unsafe} from something safe. 
The Fundamental Theorem also serves as a sanity check of the
logical relation as it guarantees that a large class of capabilities
actually inhabits it.

@@

%\bibliographystyle{IEEEtran}
\bibliographystyle{plain}
\bibliography{IEEEabrv,references}
\end{document}
\grid
